---
title: "Solow Growth Model (Part 1) — Reproducible Econometrics"
author: "Davide Yang"
output:
  html_document:
    theme: readable
    toc: true
    toc_depth: 2
    number_sections: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)

suppressPackageStartupMessages({
  library(readr)
  library(dplyr)
  library(knitr)
})

# Helpers for nicer tables
fmt_num <- function(x, digits = 3) round(x, digits)
fmt_p <- function(p) {
  ifelse(is.na(p), NA,
         ifelse(p < 0.001, "<0.001", sprintf("%.3f", p)))
}

# Path helper so the report works whether it is knitted from:
# - the project root, or
# - inside the /report folder.
# Usage: p <- rel_path("results/tables/mi_main.csv")
rel_path <- function(p) {
  if (file.exists(p)) return(p)
  alt <- file.path("..", p)
  if (file.exists(alt)) return(alt)
  return(p) # fallback (will error clearly if missing)
}
```

## Research question

This report tests the qualitative Solow-style predictions using World Bank data:

- Higher savings (investment) should be associated with higher GDP per capita.
- Higher population-growth term should be associated with lower GDP per capita.

**Model (Part 1):**

\[
\log(y_{it}) = \alpha + \beta_1 \log(s_{it}) + \beta_2 \log(n_{it} + g + \delta) + \varepsilon_{it}
\]

where we set \(g+\delta = 0.05\). This avoids dropping observations simply because population growth is negative.

## Data and sample

- Source: World Bank (downloadable automatically via script).
- Sample window: 1982–2019.
- Extremely sparse countries are excluded (high missingness in key variables).

Missing values are handled using **multiple imputation (mice)**. Results are compared to a **complete-case** regression with **HC3 robust standard errors**.

## Results

### Main (Multiple Imputation pooled)

```{r}
mi <- read_csv(rel_path("results/tables/mi_main.csv"), show_col_types = FALSE)

mi_out <- mi %>%
  transmute(
    term,
    estimate  = fmt_num(estimate),
    std_error = fmt_num(std.error),
    statistic = fmt_num(statistic),
    df        = fmt_num(df),
    p_value   = fmt_p(p.value),
    ci_low    = fmt_num(`2.5 %`),
    ci_high   = fmt_num(`97.5 %`)
  )

kable(
  mi_out,
  caption = "Main regression (MI pooled). Log-log coefficients can be read as elasticities."
)
```

**Interpretation (elasticities):**

```{r}
b_s   <- mi$estimate[mi$term == "log_s"]
b_ngd <- mi$estimate[mi$term == "log_ngd"]

cat(paste0(
  "- `log_s` is positive: a 1% increase in savings is associated with about ",
  fmt_num(b_s, 3), "% higher GDP per capita (holding the population-growth term fixed).\n",
  "- `log_ngd` is negative: a 1% increase in (n+g+δ) is associated with about ",
  fmt_num(b_ngd, 3), "% lower GDP per capita (holding savings fixed).\n"
))
```

Because this is a pooled specification, these should be treated as **associations** (so not actual causal effects).

### Robustness (Complete-case + HC3)

```{r}
cc <- read_csv(rel_path("results/tables/complete_case_hc3.csv"), show_col_types = FALSE)

cc_out <- cc %>%
  transmute(
    term,
    estimate  = fmt_num(estimate),
    std_error = fmt_num(std_error),
    statistic = fmt_num(statistic),
    p_value   = fmt_p(p_value)
  )

kable(
  cc_out,
  caption = "Robustness: complete-case OLS with HC3 robust standard errors."
)
```

The signs and magnitudes are broadly consistent across MI and complete-case, which supports the stability of our main qualitative conclusion.

## Diagnostics

```{r, out.width="85%"}
# IMPORTANT: Use an ABSOLUTE path so Pandoc can always find the image
# even when the HTML is created inside the /report folder.
img_rel <- rel_path("results/figures/resid_vs_fitted.png")
img_abs <- normalizePath(img_rel, winslash = "/", mustWork = TRUE)

knitr::include_graphics(img_abs)
```

- Residuals are mostly centered around zero.
- Outliers exist (expected in cross-country macro data).
- Heteroskedasticity is plausible → reporting **robust SE** is appropriate (as done above).

## Conclusion (Part 1)

**Main conclusion:** The data support the standard Solow-style qualitative predictions:

- Savings ↗ GDP per capita.

- Population-growth term ↘ GDP per capita.

**what's done in Part 1:**

- Fully reproducible pipeline (download → clean → impute → estimate → export).

- MI results + complete-case robust-SE check.

## Notes / limitations (what will be in Part 2)

- Panel structure is not yet modeled (country FE / year FE planned).
- Clustered standard errors by country planned.
- More robustness planned (PPP GDPpc outcome, subperiods, winsorization/outlier robustness).
